\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-15 analysis.Rmd
more stats

df |> 
  group_by(Participant, Belief) |> 
  summarize(n = n() / 108, 
            Confidence = mean(Belief_Confidence)) |> 
  pivot_wider(values_from=c("n", "Confidence"), names_from="Belief") |> 
  ungroup() |> 
  describe_posterior(centrality = "mean", test=FALSE)
```

@@

preds <- data.frame()
dats <- data.frame()
for (x in IVs) {
  for (y in IVs) {
    if (x == y) next
    print(paste(y, "~", x))
    model <- glmmTMB::glmmTMB(as.formula(
      paste(y, "~", x, "* Sex * Stimulus_Interest + (1|Participant) + (1|Stimulus)")
    ),
    data = df,
    family = glmmTMB::beta_family()
    )

@@

```{r message=FALSE, warning=FALSE, fig.width=fig.width, fig.height=fig.width}
rez_at <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Attractive, 2) + (1|Participant) + (1|Stimulus)",
  var = "Attractive", fill = "#F44336"
)
rez_gl <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Beauty, 2) + Trustworthy + (1|Participant) + (1|Stimulus)",
  var = "Beauty", fill = "#E91E63"
)
rez_tr <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Trustworthy, 2) + Beauty + (1|Participant) + (1|Stimulus)",
  var = "Trustworthy", fill = "#4CAF50"
)
rez_fa <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Familiar, 2) + (1|Participant) + (1|Stimulus)",
  var = "Familiar", fill = "#2196F3"
)

>>>>

df |> 
  group_by(Participant, Belief) |> 
  summarize(n = n() / 108, 
            Confidence = mean(Belief_Confidence)) |> 
  pivot_wider(values_from=c("n", "Confidence"), names_from="Belief") |> 
  ungroup() |> 
  describe_posterior(centrality = "mean", test=FALSE)

m <- glmmTMB::glmmTMB(Belief ~ 1 + (1|Participant) + (1|Stimulus), data=df, family="binomial")
icc(m, by_group = TRUE)
```

@@

correlation::correlation(df[IVs], partial=TRUE)
```

```{r message=FALSE, warning=FALSE}
preds <- data.frame()
dats <- data.frame()
for (x in IVs) {
  for (y in IVs) {
    if (x == y) next
    print(paste(y, "~", x))
    model <- glmmTMB::glmmTMB(as.formula(
      paste(y, "~", x, "* Sex * Stimulus_Interest + (1|Participant) + (1|Stimulus)")
    ),
    data = df,
    family = glmmTMB::beta_family()
    )

@@

```{r message=FALSE, warning=FALSE, fig.width=fig.width, fig.height=fig.width}
rez_at <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Attractive, 2) + (1|Participant) + (1|Stimulus)",
  var = "Attractive", fill = "#F44336"
)
rez_gl <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Beauty, 2) + Trustworthy + Familiar + (1|Participant) + (1|Stimulus)",
  var = "Beauty", fill = "#E91E63"
)
rez_tr <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Trustworthy, 2) + Beauty + Familiar + (1|Participant) + (1|Stimulus)",
  var = "Trustworthy", fill = "#4CAF50"
)
rez_fa <- make_model(filter(df, Stimulus_Interest == TRUE),
  formula = "Sex / poly(Familiar, 2) + Beauty + Trustworthy + (1|Participant) + (1|Stimulus)",
  var = "Familiar", fill = "#2196F3"
)

### 2022-09-15 submission/manuscript.Rmd
Update manuscript.Rmd
more stats

The final sample included 100 participants (Mean age = 27.9, SD = 8.5, range: [19, 66]; Sex: 48.0% females, 52.0% males, 0.0% other).

@@

The real-fake ratings were converted into two scores.

@@

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle].

@@

Only one image file, classified as fake in 87.4\% of trials, yielded a strong simulation monitoring bias (> 85\%). This image was removed from further analysis (leaving 108 trials per participant). On average, across participants, 44\% of trials (95\% CI [0.11, 0.64]) was considered fake and 56\% of trials (95\% CI [0.36, 0.89]) was considered real.

There was a significant negative effect of the delay of re-exposure (with 95\% of values between 1.58 and 30.31 min), suggesting that shorter delays were associated with a slight bias towards the belief of reality (60\% at a theoretical delay of 0), which decreased to 50\% at a theoretical delay of 60 min ($OR = 0.99$, $95\% CI = [0.99, 1.00]$ $z = -2.27$, $p = 0.023$). There was also a significant negative effect on judgment confidence, but only in the real condition ($\beta = -0.005$, $95\% CI = [-0.1, 0.0]$, $p = .023$).

### Determinants of Simualtion Monitoring

@@

Although expected, the first striking result is that despite all items being real photos from the same database, all people easily believe (to high degrees of confidence) that a significant proportion of them was fake. This is a testimony to the existing expectations regarding CGI technology, as much as to the volatility of our sense of reality.

>>>>

The final sample included 100 participants (Mean age = 27.9, SD = 8.5, range: [19, 66]; Sex: 48% females, 52% males).

@@

The real-fake ratings (measured originally on a [-1, 1] analog scale) were converted into two scores, corresponding to two distinct mechanisms: the dichotomous *belief* (real or fake, derived based on the sign of the rating) and the *confidence* (the rating's absolute value) associated with that belief. Models predicting the former were set as logistic mixed models (with the participants and images entered as random factors), and models modeling the latter, as well as the other face ratings (attractiveness, beauty, trustworthiness and familiarity) were modeled using beta regression models (suited for an outcome variable expressed in percentages).

We started by investigating the effect of the procedure and instructions to check whether the stimuli (which were real pictures of faces) were indeed judged as fake in a sufficient proportion to warrant their analysis. Additionally, we assessed the effect of the re-exposure delay, i.e., the time between the first presentation of the image (corresponding to the face ratings) and the second presentation (for the real-fake rating).

The determinants of reality beliefs were modeled separately for attractiveness, beauty, trustworthiness, and familiarity, using second order raw polynomials coefficients to allow for possible quadratic relationships (**Figure 2**. Aside from attractiveness (conceptualized as a general construct), models for beauty, trustworthiness and familiarity were adjusted for the the two remaining variables *mutatis mutandis*. We took into account the gender of participants and stimuli by retaining the pictures that were aligned with the participants' sexual preference (e.g., female faces for homosexual females, male faces for heterosexual females, and both for bisexual participants), and modeling the interaction with the participants' gender.

The analysis was carried out using *R 4.2* [@RCoreTeam2022], the *tidyverse* [@wickham2019], and the *easystats* collection of packages [@bayestestRArticle; @correlationArticle; @performanceArticle; @insightArticle]. As all the details, scripts and complimentary analyses are available in open-access, the manuscript will focus on significant results.

@@

@@

Only one image file yielded a strong simulation monitoring bias (> 85\%), being classified as fake in 87.4\% of trials. This image was removed from further analysis, leaving 108 trials per participant. On average, across participants, 44\% of images (95\% CI [0.11, 0.64]) were judged as fake and 56\% of images (95\% CI [0.36, 0.89]) as real. An intercept-only model with the participants and images as random factors showed that the Intraclass Correlation Coefficient (ICC), which can be interpreted as the proportion of variance explained by the random factors, was of 10.5% for the participants and 8.7% for the pictures.

There was a significant negative effect of the delay of re-exposure (with 95\% of values between 1.58 and 30.31 min), suggesting that shorter delays were associated with a slight bias towards the belief of reality (60\% at a theoretical delay of 0), which decreased to 50\% at a theoretical delay of 60 min ($OR = 0.99$, $95\% CI = [0.99, 1.00]$ $z = -2.27$, $p = .023$). There was also a significant negative effect on judgment confidence, but only in the real condition ($\beta = -0.005$, $95\% CI = [-0.1, 0.0]$, $p = .023$).

### Determinants of Simulation Monitoring

Attractiveness had a significant positive and linear relationship ($R^2_{marginal}$ = 2.8%) with the belief that a stimulus was real ($\beta_{poly1} = 16.37$, $95\% CI = [7.76, 24.98]$ $z = 3.73$, $p < .001$) for males, and a quadratic relationship for females ($\beta_{poly2} = 7.77$, $95\% CI = [1.41, 14.13]$ $z = 2.40$, $p = .017$), with both non-attractive and attractive faces being judged as more real. No significant relationship was found between attractiveness ratings and belief confidence, aside of a similar trend for females only, for faces judged as real ($\beta_{poly2} = 4.38$, $95\% CI = [0.96, 7.79]$ $z = 2.51$, $p = .012$).

Beauty, adjusted for trustworthiness and familiarity, had a significant positive and linear relationship ($R^2_{marginal}$ = 3.5%) with the belief that a stimulus was real ($\beta_{poly1} = 9.54$, $95\% CI = [1.43, 17.65]$ $z = 2.31$, $p = .021$) for males only. No effect on confidence was found, aside from a quadratic relationship for females for faces judged as fake, suggesting that non-beautiful and highly beautiful faces were rated as fake with more confidence than average faces ($\beta_{poly2} = 6.61$, $95\% CI = [1.98, 11.24]$ $z = 2.80$, $p = .005$).

Trustworthiness, adjusted for beauty and familiarity, had a significant positive and linear relationship ($R^2_{marginal}$ = 3.0%) with the belief that a stimulus was real ($\beta_{poly1} = 11.60$, $95\% CI = [4.15, 19.06]$ $z = 3.05$, $p = .002$) for females only. No effect on confidence was found, aside from a quadratic relationship for females for faces judged as real, suggesting that non-trustworthy and highly trustworthy faces were rated as real with more confidence than average faces ($\beta_{poly2} = 6.47$, $95\% CI = [1.73, 11.21]$ $z = 2.68$, $p = .007$).

We did not find any significant relationships for familiarity adjusted for beauty and trustworthiness ($R^2_{marginal}$ = 3.0%). However, a significant positive and linear relationship was found with the confidence in faces judged as real ($\beta_{poly1} = 9.31$, $95\% CI = [3.45, 15.17]$ $z = -3.11$, $p = .002$), and a quadratic relationship for faces judged as fake ($\beta_{poly1} = -12.67$, $95\% CI = [-19.87, -5.47]$ $z = -3.45$, $p < .001$; $\beta_{poly2} = 8.14$, $95\% CI = [0.01, 16.28]$ $z = 1.96$, $p = .05$), for males only, suggesting that faces judged as real with more confidence when they are familiar, and judged as fake with less confidence when they are of not familiar or highly familiar.

@@

Although expected, the first striking result is that despite all items being real photos from the same database, all people easily believe (to high degrees of confidence) that a significant proportion of them was fake. This is a testimony to the existing expectations regarding CGI technology, as much as to the volatility of our sense of reality. In fact, stimuli-related and participant-related characteristics accounted for about only less than 20% of the variance in the beliefs, suggesting that a large part of it is related to other subjective processes.

Though attractiveness does not seem to be the primary drive underlying simulation monitoring of face images, it does nonetheless display an association.
